Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-12-24 paper/paper.tex
update

\begin{abstract}
We introduce a method of using generic programming tensor network to compute various properties of independent sets.
This method is versertile such that by engineering tensor element algebra, the type of graph problems that can be computed using this method includes: independence number, the number of independent sets, independence polynomial, maximal independence polynomial, and both counting and enumeration of all maximum independence sets, the large-size independent sets, and all maximal independent sets.
%Although the computational complexity inevitably scales exponentially with treewidth of the graph,
%our method using tensor network contraction with generic programming is highly versatile and is able to compute properties that are otherwise not known to be possible for up to several hundred vertices.
At the same time, our algorithms are very simple to implement and one can directly utilize recent advances in tensor network contraction techniques like near optimal contraction order finding and slicing.
One can easily find a good contraction order for the tensor network representation of a graph $G$ with thousands of vertices. This contraction order gives tensor network contraction complexity $2^{{\rm tw}(G)}$, with ${\rm tw} (G)$ the tree width of the problem graph.
To demonstrate the versatility of this tool, we apply it to a few examples including the calculations of the hard-square entropy constant, Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs\blue{remove some examples}.
\end{abstract}

@@

%more concretely, an maximal independent set of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
%and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
%In this paper, the term ``maximal'' has a different meaning from ``maximum''  which we discussed above; a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.
In this paper, we do not limit our discussion to finding one of the maximum indepedent sets and its size.
There are many other interesting problems pertaining to independent sets, such as the number of independent sets,
independence polynomial, maximal independence polynomial, and enumeration of MIS or maximal independent sets.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready};
they can, for example, be used to understand phase transitions, to analyse the overlap gap properties~\cite{Gamarnik2013, Gamarnik2019} and to identify harder graphs in an ensemble of graphs \red{cite experiment}.
The problem of computing the independence polynomial belongs to the complexity class \#P-hard. One approach to exactly compute the independence polynomial requires a computational time of $O(1.442^{|V|})$~\cite{Ferrin2014}
\blue{I am not sure about this complexity, this is based on the naive analysis of theorem 2.2 in ~\cite{Ferrin2014}}. There are some interesting works on efficiently approximating the independence polynomial~\cite{Harvey2018}, but in this work, we focus on computing the independence polynomial exactly.
We introduce a tensor-network based framework to compute the various properties pertaining to independent sets.
We map the independent set problem into generic tensor network contraction with specially designed tensor element algebra.
Its algorithmic complexity is similar to that of dynamic programming, i.e. scales exponentially as ${\rm tw}(G)$.
%But the tensor network approach is highly versatile so different problems can be generically solved with minimal implementation efforts without loss of efficiency.
We benchmark our algorithms by computing various independent set properties on central processing units (CPUs) and graphics processing units (GPUs) and show its good performance on sparse graphs.
Its high performance benefits from recent advances in random tensor-network contraction for the purposes of quantum circuit simulations~\cite{Gray2021, Pan2021, Kalachev2021}.
People developed algorithms to optimize the contraction order for tensor networks with thousands of vertices,
and the optimized contraction order gives us a near optimal space complexity $2^{\sim{\rm tw}(G)}$ to contract the tensor network.
For cases when the tensors being too large to fit into a GPU, slicing technique can be used to further reduce the space consumption.
Lastly, we provide a few examples and show that this toolbox can be used to calculate the hard-square entropy constant,
Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs. 
The implementation of our algorithm is publicly available online: ~\cite{repolink}.

@@

This motivates us to think about what is the most general element type that is allowed in a tensor network contraction program.
We find that as long as the algebra of tensor elements forms a commutative semiring, the tensor network contraction result will be valid and be independent of the contraction order.
A commutative semiring is a semiring with its multiplication operation being commutative, and a semiring is a ring without additive inverse.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for any arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order. In the following sections, we show how to compute the independence polynomial,
the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS and maximal independent sets of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

@@

\section{Independence polynomial and maximal independence polynomial}
\subsection{Independence polynomial}\label{sec:indpoly}
The independence polynomial is an important graph polynomial that contains the counting information of independent sets. It is defined as
\begin{equation}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$. The total number of independent sets is thus equal to $I(G, 1)$.
To compute the independence polynomial of a graph $G = (V, E)$, we reinterpret this problem as a tensor network contraction problem.
We map a vertex $i\in V$ to a label $s_i \in \{0, 1\}$ of dimension $2$ in a tensor network, where we use $0$ ($1$) to denote a vertex absent (exists) in the set.
For each label $s_i$, we defined a prametrized rank-one vertex tensor $W(x_i)$ indexed by it as
\begin{equation}
    W(x_i) = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right).
\end{equation}
On each edge $(u, v)$, we define a matrix $B$ indexed by $(s_u, s_v)$ as
\begin{equation}
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation}
The corresponding tensor network contraction gives
\begin{equation}\label{eq:idp}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$.
In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial.
The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise.

@@

\subsection{Maximal independence polynomial}
Sometimes, one may be interested in knowing maximal solutions to understand why his or her program is trapped in a local minimum.
Then instead of counting all independent sets, the maximal independence polynomial counts the number of maximal independent sets of various sizes~\cite{Hu2017} would be more helpful.
 Concretely, it is defined as
\begin{equation}
I_{\rm max}(G, x) = \sum_{k=0}^{\alpha(G)} b_k x^k,
\end{equation}
where $b_k$ is the number of maximal independent sets of size $k$ in $G$.
Obviously, $b_{k} \leq a_{k}$ and $b_{\alpha(G)} = a_{\alpha(G)}$. $I_{\rm max}(G, 1)$ counts the total number of maximal independent sets~\cite{Gaspers2012, Manne2013}, where the fastest algorithm currently has a runtime of $O(1.3642^{|V|})$~\cite{Gaspers2012}.
If we want to find an MIS, $b_{k}$ counts the number of local optimum at size $k < \alpha(G)$, and can, in some cases, provide hints on the difficulty of finding the MIS using local algorithms~\red{cite experiment}.
The uni-modality, log-concavity, and real-rootness properties of the maximal independence polynomial for special classes of graphs have also been studied~\cite{Hu2017}.

@@

\section{Enumeration of configurations and enumeration bounds}
\subsection{Enumeration of independent sets}
The enumeration problems of independent sets are also interesting and has been studied extensively in the literature~\cite{Bron1973, Eppstein2010, Johnson1988}, including,
for example, the enumeration of all independent sets, the enumeration of all maximal independent sets, or the enumeration of all MISs.
The standard algorithm to enumerate all maximal independent sets is the Bron-Kerbosch algorithm~\cite{Bron1973}, which, of course, includes the enumeration of all MISs and can be used to enumerate all independent sets as well.
Here, we adapt the tensor element types for different enumeration problems.
For example, our tensor-network based algorithm to enumerate all MISs is expectedly much faster and more space-efficient than the Bron-Kerbosch algorithm, which lists all maximal independent sets.

To enumerate all independent sets, we input a set of configurations into the tensor elements. More concretely, we design a new element type having the following algebra
\begin{equation}
\eqname{SN}
\begin{split}
    s \oplus t &= s \cup t\\
    s \odot t &= \{\sigma \lor^\circ \tau \, | \, \sigma \in s, \tau \in t\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}.
\end{split}\label{eq:set}
\end{equation}
where $s$ and $t$ are each a set of $|V|$-bit strings and $\lor^\circ$ is the bitwise OR operation over two bit strings.
\begin{example}\label{eg:setalgebra}
    For elements being bit vectors of length $5$, we have the following set algebra
\begin{equation*}
\begin{split}
    &\{00001\} \oplus \{01110, 01000\} = \{01110, 01000\} \oplus \{00001\} = \{00001,01110, 01000\}\\
    &\{00001\} \oplus \{\} = \{00001\}\\
&\\
    &\{00001\} \odot \{01110, 01000\} = \{01110, 01000\} \odot \{00001\} = \{01111, 01001\}\\
    &\{00001\} \odot \{\} = \{\}\\
    &\{00001\} \odot \{00000\} = \{00001\}
\end{split}
\end{equation*}
\end{example}

@@

\subsection{Bounding the MIS enumeration space}
When we use the algebra in \Eq{eq:countingtropicalset} to enumerate all MIS configurations, we find that the program stores significantly more intermediate configurations than necessary and thus incur significant overheads in space.
To speed up the computation and reduce space overhead, we use $\alpha(G)$ to bound the searching space.
First, we compute the value of $\alpha(G)$ with tropical algebra and cache all intermediate tensors.
Then, we compute a boolean mask for each cached tensor, where we use a boolean true to represent a tensor element having a contribution to the MIS (i.e.\ with a non-zero gradient) and boolean false otherwise.
Finally, we perform masked matrix multiplication using the new element type with the above algebra, \Eq{eq:countingtropicalset}, for obtaining all configurations.
Note that these masks in fact correspond to tensor elements with non-zero gradients with respect to the MIS size; we compute these masks by back propagation of the gradients.
To derive the back-propagation rule for tensor contraction,
we first reduce the problem to finding the back-propagation rule of a tropical matrix multiplication $C = A B$.
Since $ C_{ik} = \bigoplus_{j} \ A_{ij} \odot B_{jk} = \max_{j} \ A_{ij} \odot B_{jk}$ with tropical algebra, we have the following inequality
\begin{equation}
    A_{ij} \odot B_{jk} \leq C_{ik}.
\end{equation}
Here $\leq$ on tropical numbers are the same as the real-number algebra.
The equality holds for some $j'$, which means $A_{ij'}$ and $B_{j'k}$ have contributions to $C_{ik}$.
Since $A_{ij} \odot B_{jk} = A_{ij} + B_{jk}$, one can move $B_{jk}$ to the right hand side of the inequality: 
\begin{equation}
    A_{ij} \leq C_{ik} \odot B_{jk}^{\circ -1}
\end{equation}
where ${}^{\circ -1}$ is the element-wise multiplicative inverse on tropical algebra (which is the additive inverse on real numbers).
The inequality still holds if we take the minimum over $k$: 
\begin{equation}
    A_{ij} \leq \min_{k}(C_{ik} \odot B_{jk}^{\circ -1}) = \left(\max_{k} \left(C_{ik}^{\circ -1} \odot B_{jk} \right) \right)^{\circ -1} = \left(\bigoplus_{k} \left(C_{ik}^{\circ -1} \odot B_{jk} \right) \right)^{\circ -1} = \left( C^{\circ-1} B^{\mathsf{T}} \right)^{\circ -1}_{ij}.
\end{equation}
On the right hand side, we transform the operation into a tropical matrix multiplication so that we can utilize the fast tropical BLAS routines~\cite{TropicalGEMM}.
Again, the equality holds if and only if the element $A_{ij}$ has a contribution to $C$ (i.e.\ having a non-zero gradient).
Let the gradient mask for $C$ be $\overline C$; the back-propagation rule for gradient masks reads
\begin{equation}
\overline{A}_{ij} = \delta \left(A_{ij}, \left( \left( C^{\circ-1} \circ \overline C \right) B^{\mathsf{T}} \right)_{ij}^{\circ -1} \right),
\end{equation}
where $\circ$ is the element-wise product, boolean false is treated as the tropical number $\mymathbb{0}$, and boolean true is treated as the tropical number $\mymathbb{1}$.
This rule defined on matrix multiplication can be easily generalized to tensor contraction by replacing the matrix multiplication between $C^{\circ-1} \circ \overline C$ and $B^{\mathsf{T}}$ by a tensor contraction.
With the above method, one can significantly reduce the space needed to store the intermediate configurations.

@@

We also benchmarked our method with the Bron-Kerbosch method to enumerate and count maximal independent sets in \Fig{fig:benchmark-maximal}.
Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
The implementation of Bron-Kerbosch algorithm is from the package LightGraphs~\cite{LightGraphs},
due to the memory limit, this algorithm stops working at size $70$ and above.

@@

\section{Discussion and conclusion}
We have introduced a new approach based on tensor network contraction to compute the independence number, independence polynomial, maximal independence polynomial, and enumeration of MIS configurations.
We derived the backward rule for tropical tensor network to bound the search of solution space.
Although many of these properties are global, we can encode them to different tensor element types as commutative semirings.
The power of our tensor network approach is not only limited to the independent set problem, in \App{app:otherproblems}, we show how to map the matching problem, $k$-coloring problem, max cut problem and set packing problem to tensor networks.
The independence polynomial is closely related to the matching polynomial, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013}.
In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{Levit2005}.
Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
We hope our algorithms to calculate the independence polynomial can also help further research in these directions.
We show some of the Julia language implementations in Appendix~\ref{sec:technical} and you will find it surprisingly short.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}.
%What we need to do is just defining two operations $\oplus$ and $\odot$ and two special elements $\mymathbb{0}$ and $\mymathbb{1}$.
%The style that we program is called generic programming,

>>>>

\begin{abstract}
We introduce a method of using generic programming tensor network to compute various properties of independent sets,
which includes: maximum size of such sets, the number of sets of a given size, enumeration of sets of a given size.
It can be generalized to compute properties of other problems like cut size, coloring, maximal cliques et al.
%Although the computational complexity inevitably scales exponentially with treewidth of the graph,
%our method using tensor network contraction with generic programming is highly versatile and is able to compute properties that are otherwise not known to be possible for up to several hundred vertices.
At the same time, our algorithms are very simple to implement and one can directly utilize recent advances in tensor network contraction techniques like near optimal contraction order finding and slicing.
%One can easily find a good contraction order for the tensor network representation of a graph $G$ with thousands of vertices. 
The algorithmic complexity of this approach is $2^{{\rm tw}(G)}$, where ${\rm tw} (G)$ the tree width of the problem graph.
To demonstrate the versatility of this tool, we apply it to a few examples including the calculations of the hard-square entropy constant, Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs\blue{remove some examples}.
\end{abstract}

@@

%more concretely, an maximal independent set of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
%and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
%In this paper, the term ``maximal'' has a different meaning from ``maximum''  which we discussed above; a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.
In this paper, we do not limit our discussion to finding one of the maximum indepedent sets and its size.
There are many other interesting problems pertaining to independent sets, such as the number of independent sets at a given size
and enumerating independent sets at a give size or what we call computing \textit{properties}.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready};
they can, for example, be used to understand phase transitions, to analyse the overlap gap properties~\cite{Gamarnik2013, Gamarnik2019} and to identify harder graphs in an ensemble of graphs \red{cite experiment}.
The problem of computing the independence polynomial belongs to the complexity class \#P-hard. One approach to exactly compute the independence polynomial requires a computational time of $O(1.442^{|V|})$~\cite{Ferrin2014}
\blue{I am not sure about this complexity, this is based on the naive analysis of theorem 2.2 in ~\cite{Ferrin2014}}. There are some interesting works on efficiently approximating the independence polynomial~\cite{Harvey2018}, but in this work, we focus on computing the independence polynomial exactly.
We introduce a tensor-network based framework to compute the various properties pertaining to independent sets.
We map the independent set problem into generic tensor network contraction with specially designed tensor element algebra.
Its algorithmic complexity is similar to that of dynamic programming, i.e. scales exponentially as ${\rm tw}(G)$.
%But the tensor network approach is highly versatile so different problems can be generically solved with minimal implementation efforts without loss of efficiency.
We benchmark our algorithms by computing various independent set properties on central processing units (CPUs) and graphics processing units (GPUs) and show its good performance on sparse graphs.
Its high performance benefits from recent advances in random tensor-network contraction for the purposes of quantum circuit simulations~\cite{Gray2021, Pan2021, Kalachev2021}.
People developed algorithms to optimize the contraction order for tensor networks with thousands of vertices,
and the optimized contraction order gives us a near optimal space complexity $2^{\sim{\rm tw}(G)}$ to contract the tensor network.
For cases when the tensors being too large to fit into a GPU, slicing technique can be used to further reduce the space consumption.
Lastly, we provide a few examples and show that this toolbox can be used to calculate the hard-square entropy constant,
Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs. 
The implementation of our algorithm is publicly available online: ~\cite{repolink}.

@@

This motivates us to think about what is the most general element type that is allowed in a tensor network contraction program.
We find that as long as the algebra of tensor elements forms a commutative semiring, the tensor network contraction result will be valid and be independent of the contraction order.
A commutative semiring is a semiring with its multiplication operation being commutative, and a semiring is a ring without additive inverse.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for any arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order.
In the following sections, we show how to compute the properties of independent sets by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the properties that can be solved by various tensor element types.

@@

\section{Independence polynomial}
\subsection{Independence polynomial}\label{sec:indpoly}
The independence polynomial is an important graph polynomial that contains the counting information of independent sets. It is defined as
\begin{equation}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in $G$. The total number of independent sets is thus equal to $I(G, 1)$.
To compute the independence polynomial of a graph $G = (V, E)$, we reinterpret this problem as a tensor network contraction problem.
We map a vertex $i\in V$ to a label $s_i \in \{0, 1\}$ of dimension $2$ in a tensor network, where we use $0$ ($1$) to denote a vertex absent (exists) in the set.
For each label $s_i$, we defined a prametrized rank-one vertex tensor $W(x_i)$ indexed by it as
\begin{equation}
    W(x_i) = \left(\begin{matrix}
        1 \\
        x_i
    \end{matrix}\right).
\end{equation}
On each edge $(u, v)$, we define a matrix $B$ indexed by $(s_u, s_v)$ as
\begin{equation}
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right), \label{eq:tensor}
\end{equation}
The corresponding tensor network contraction gives
\begin{equation}\label{eq:idp}
    P(G, \{x_1, x_{2}, \ldots,x_{|V|}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$.
In the special case of $x_i = x$, the contraction result directly corresponds to the independence polynomial.
The connection can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise.

@@

@@

\section{Enumeration of independent sets}
\subsection{Set algebra for configuration enumeration}
The enumeration problems of independent sets are also interesting and has been studied extensively in the literature~\cite{Bron1973, Eppstein2010, Johnson1988}, including,
for example, the enumeration of all independent sets, the enumeration of all maximal independent sets, or the enumeration of all MISs.
Here, the term ``maximal'' has a different meaning from ``maximum''  which we discussed above;
a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.
We can also compute the maximal independence polynomial and enumerate maximal independent sets with a different tensor network mapping, and this is described in \Sec{sec:maximal}.
To enumerate all independent sets, we designed an algebra defined on sets of bitstrings.
\begin{equation}
\eqname{SN}
\begin{split}
    s \oplus t &= s \cup t\\
    s \odot t &= \{\sigma \lor^\circ \tau \, | \, \sigma \in s, \tau \in t\}\\
    \mymathbb{0} &= \{\}\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}.
\end{split}\label{eq:set}
\end{equation}
where $s$ and $t$ are each a set of $|V|$-bit strings and $\lor^\circ$ is the bitwise OR operation over two bit strings.
\begin{example}\label{eg:setalgebra}
    For elements being bit vectors of length $5$, we have the following set algebra
\begin{equation*}
\begin{split}
    &\{00001\} \oplus \{01110, 01000\} = \{01110, 01000\} \oplus \{00001\} = \{00001,01110, 01000\}\\
    &\{00001\} \oplus \{\} = \{00001\}\\
&\\
    &\{00001\} \odot \{01110, 01000\} = \{01110, 01000\} \odot \{00001\} = \{01111, 01001\}\\
    &\{00001\} \odot \{\} = \{\}\\
    &\{00001\} \odot \{00000\} = \{00001\}
\end{split}
\end{equation*}
\end{example}

@@

\subsection{Bounding the MIS enumeration space}
When we use the algebra in \Eq{eq:countingtropicalset} to enumerate all MIS configurations, we find that the program stores significantly more intermediate configurations than necessary and thus incur significant overheads in space.
To speed up the computation and reduce space overhead, we bound the searching space using the MIS size $\alpha(G)$.
First, we compute the value of $\alpha(G)$ with tropical algebra and cache all intermediate tensors.
Then, we compute a boolean mask for each cached tensor, where we use a boolean true to represent a tensor element having a contribution to the MIS (i.e.\ with a non-zero gradient) and boolean false otherwise.
Finally, we perform masked matrix multiplication using the new element type with the above algebra, \Eq{eq:countingtropicalset}, for obtaining all configurations.
Note that these masks in fact correspond to tensor elements with non-zero gradients with respect to the MIS size; we compute these masks by back propagating the gradients.
To derive the back-propagation rule for tropical tensor contraction,
we first reduce the problem to finding the back-propagation rule of a tropical matrix multiplication $C = A B$.
Since $ C_{ik} = \bigoplus_{j} \ A_{ij} \odot B_{jk} = \max_{j} \ A_{ij} \odot B_{jk}$ with tropical algebra, we have the following inequality
\begin{equation}
    A_{ij} \odot B_{jk} \leq C_{ik}.
\end{equation}
Here $\leq$ on tropical numbers are the same as the real-number algebra.
The equality holds for some $j'$, which means $A_{ij'}$ and $B_{j'k}$ have contributions to $C_{ik}$.
Intuitively, one can use this relation to indentify elements with nonzero gradients in $A$ and $B$ directly,
however, doing this might make us lost the advantage of using BLAS libraries~\cite{TropicalGEMM}.
Since $A_{ij} \odot B_{jk} = A_{ij} + B_{jk}$, one can move $B_{jk}$ to the right hand side of the inequality: 
\begin{equation}
    A_{ij} \leq C_{ik} \odot B_{jk}^{\circ -1}
\end{equation}
where ${}^{\circ -1}$ is the element-wise multiplicative inverse on tropical algebra (which is the additive inverse on real numbers).
The inequality still holds if we take the minimum over $k$: 
\begin{equation}
    A_{ij} \leq \min_{k}(C_{ik} \odot B_{jk}^{\circ -1}) = \left(\max_{k} \left(C_{ik}^{\circ -1} \odot B_{jk} \right) \right)^{\circ -1} = \left(\bigoplus_{k} \left(C_{ik}^{\circ -1} \odot B_{jk} \right) \right)^{\circ -1} = \left( C^{\circ-1} B^{\mathsf{T}} \right)^{\circ -1}_{ij}.
\end{equation}
On the right hand side, we transform the operation into a tropical matrix multiplication so that we can utilize the fast tropical BLAS routines~\cite{TropicalGEMM}.
Again, the equality holds if and only if the element $A_{ij}$ has a contribution to $C$ (i.e.\ having a non-zero gradient).
Let the gradient mask for $C$ be $\overline C$; the back-propagation rule for gradient masks reads
\begin{equation}
\overline{A}_{ij} = \delta \left(A_{ij}, \left( \left( C^{\circ-1} \circ \overline C \right) B^{\mathsf{T}} \right)_{ij}^{\circ -1} \right),
\end{equation}
where $\circ$ is the element-wise product, boolean false is treated as the tropical number $\mymathbb{0}$, and boolean true is treated as the tropical number $\mymathbb{1}$.
This rule defined on matrix multiplication can be easily generalized to tensor contraction by replacing the matrix multiplication between $C^{\circ-1} \circ \overline C$ and $B^{\mathsf{T}}$ by a tensor contraction.
With the above method, one can significantly reduce the space needed to store the intermediate configurations by setting the tensor elements masked false to zero during contraction.

@@

The standard algorithm to enumerate all maximal independent sets is the Bron-Kerbosch algorithm~\cite{Bron1973}, which, of course, includes the enumeration of all MISs and can be used to enumerate all independent sets as well.
We also benchmarked our method with the Bron-Kerbosch method to enumerate and count maximal independent sets in \Fig{fig:benchmark-maximal}.
Since the network for computing maximal independent set properties is different from the one for computing the independence polynomial,
the tree width of this tensor network is significantly larger, hence only small graphs can be benchmarked.
The time for the tensor netwrok approach and the Bron-Kerbosch approach to enumerate all maximal independent sets are comparable,
while the tensor network does counting much more efficiently.
The implementation of Bron-Kerbosch algorithm is from the package LightGraphs~\cite{LightGraphs},
due to the memory limit, this algorithm stops working at size $70$ and above.

@@

\section{Discussion and conclusion}
We have introduced a new approach based on tensor network contraction to compute various properties of independent sets,
including the independence number, number of independent set of a given size, and enumeration of independent sets of a give size.
We derived the backward rule for tropical tensor network to bound the search of solution space.
Although many of these properties are global, we can encode them to different tensor element types as commutative semirings.
The power of our tensor network approach is not only limited to the independent set problem, in \App{app:otherproblems}, we show how to map the matching problem, $k$-coloring problem, max cut problem and set packing problem to tensor networks.
The independence polynomial is closely related to the matching polynomial, the clique polynomial~\cite{Hoede1994}, and the vertex cover polynomial~\cite{Akbari2013}.
In fact, the independence polynomial can be viewed as a generalization of the matching polynomial, since the matching polynomial of a graph $G$ is the same as the independence polynomial of the line graph of $G$~\cite{Levit2005}.
Thus, our algorithm to compute the independence polynomial can also be used to compute the above graph polynomials. Some properties of the independence polynomial, such as the uni-modality, log-concavity, and roots of the polynomial, are well studied in the literature for special classes of graphs~\cite{Levit2005}.
We hope our algorithms to calculate the independence polynomial can also help further research in these directions.
We show some of the Julia language implementations in Appendix~\ref{sec:technical} and you will find it surprisingly short.
A complete implementation can be found in our Github repository~\cite{GraphTensorNetworks}.
%What we need to do is just defining two operations $\oplus$ and $\odot$ and two special elements $\mymathbb{0}$ and $\mymathbb{1}$.
%The style that we program is called generic programming,
